Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2023-02-10 write-up/fda-recurrent-supplement.tex
Fixing up the paper by removing the pmcc reference to appendix as its a side point

>>>>

\section{Measurement-error models with event processes}
\label{section:memproblems}

One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error. A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory. For example, take~$\eta_i$ to be a zero-mean Gaussian process with
\begin{equation}\label{eq:jm}
x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
\end{equation}
where~$\epsilon_i(t)$ is a white-noise measurement error term. Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
\[
\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
\]
where the expectation is a Gaussian integral, albeit infinite-dimensional.  For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}. This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process. To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.

### 2023-02-10 write-up/fda-recurrent.tex
Fixing up the paper by removing the pmcc reference to appendix as its a side point

\subsection{Measurement-error models with event processes}
\label{section:memproblems}

One potential criticism for~\eqref{eq:hazard} is that the health process may be measured with error. A common mathematical strategy for joint models is to consider an unobservable, latent process~$\eta_i$ such that~$\bfT_i \indep x_i \given \eta_i$, i.e., the two processes are conditionally independent given the latent trajectory. For example, take~$\eta_i$ to be a zero-mean Gaussian process with
\begin{equation}\label{eq:jm}
x_i(t) = \eta_i (t) + \epsilon_i (t),\quad \text{ and } \log h_i (t
\given \eta ) = \log h_0 (t) + g_t \left( H_{i,t}^N \right)^{\prime}
\alpha + \int_{t-\Delta}^t \eta_i (s) \beta (s) ds
\end{equation}
where~$\epsilon_i(t)$ is a white-noise measurement error term. Thus, $\eta_i (t)$ is the ``true and unobserved value of the longitudinal outcome'' \citep[Sec. 2.1, pp.3]{Rizopoulos2010}. The conditional survival function of the $j$th event~$T_j$ given~$\bfx_i$ and all prior events~$\bfT_{-j} := \{ T_1, \ldots, T_{j-1} \}$ is
\[
\pr \left ( T_j > t+s \Given \H_{i,\tau_i}^X, \bfT_{-j} = \bft_{-j}, T_j > t \right) = E \left( - \exp \left( \int_{t}^{t+s} h_i (u \given \eta ) du \right) \Given \bfx_i, \bfT_{-j} = \bft_{-j} \right),
\]
where the expectation is a Gaussian integral, albeit infinite-dimensional.  For most choices of covariance structure, if the white-noise error term is non-zero then the above calculation will show the conditional survival function depends not only on past $x$-values, but also on future $x$-values, i.e.,~$\bfx$ and~$\bfT$ do not satisfy \emph{independent evolution}~\citep{DempseyPMCC2}. This is quite unnatural, as~\eqref{eq:jm} suggests the instantaneous risk of an event at time~$t$ depends on future values of the sensor process. To ensure independent evolution, we set~$\epsilon (t) \equiv 0$ and treat~$\bfx_i$ as a Gaussian process measured without a white-noise error term.
Our procedure will only rely on the Gaussian assumption when~$\bfx$ is not fully observed; see Section~\ref{section:missingdata} for details on handling of missing data.

@@

\begin{table}[!th]
\begin{tabular}{c | c c c c c}
& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
\multirow{3}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
 & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\
  & P-MISE & 0.225 & 0.229 & 0.233 & 0.239 \\ \hline
\multirow{3}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
 & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\
  & P-MISE & 0.107 & 0.109 & 0.111 & 0.117 \\ \hline
\multirow{3}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
 & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\
  & P-MISE & 0.105 & 0.103 & 0.103 & 0.105 \\\hline
\multirow{3}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
 & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\
  & P-MISE & 0.273 & 0.270 & 0.269, & 0.269 \\\hline
\multirow{3}{*}{37} & MISE & 0.589 & 0.588 & 0.589 & 0.591 \\
 & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\
  & P-MISE & 0.473 & 0.470 & 0.469 & 0.470 \\\hline
\end{tabular}
\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
\label{tab:mise}
\end{table}

@@

\begin{figure}[!th]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{../figures/smoothed_eda.png}
  \caption{Electrodermal activity (EDA)}
  \label{fig:mean_eda}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{../figures/smoothed_ai.png}
  \caption{Activity Index (AI)
}  \label{fig:mean_acc}
\end{subfigure}
\caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
\label{Reduced Rank Mixed Effects Models for Spatially Correlated Hierarchical Functional Data
}
\label{fig:mean_edacc}
\end{figure}

@@

% \begin{figure}[!th]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{../figures/missing_data_eda.png}
%   \caption{Electrodermal activity (EDA)}
%   \label{fig:missing_eda}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{../figures/missing_data_ai.png}
%   \caption{Activity Index (AI)
% }  \label{fig:missing_acc}
% \end{subfigure}
% \caption{Patient average fraction missingness in 30-minute windows.}
% \label{fig:mean_edacc}
% \end{figure}

>>>>

@@

\begin{table}[!th]
\begin{tabular}{c | c c c c c}
& & \multicolumn{4}{c}{Sampling rate} \\ \cline{3-6}
$\Delta=$ & & 0.5 & 1 & 2 & 4 \\ \hline
\multirow{3}{*}{26} & MISE & 0.400 & 0.404 & 0.409 & 0.413 \\
 & Variance & $1.1 \times 10^{-2}$ & $1.3 \times 10^{-2}$ & $1.5 \times 10^{-2}$ & $1.9 \times 10^{-2}$ \\
  & P-MISE & 0.225 & 0.229 & 0.233 & 0.239 \\ \hline
\multirow{3}{*}{29} & MISE & 0.240 & 0.243 & 0.244 & 0.251 \\
 & Variance & $9.2 \times 10^{-3}$ & $1.1 \times 10^{-2}$ & $1.2 \times 10^{-2}$ & $1.8 \times 10^{-2}$ \\
  & P-MISE & 0.107 & 0.109 & 0.111 & 0.117 \\ \hline
\multirow{3}{*}{32} & MISE & 0.105 & 0.103 & 0.103 & 0.105 \\
 & Variance & $8.0 \times 10^{-3}$ & $8.8 \times 10^{-3}$ & $1.0 \times 10^{-2}$ & $1.4 \times 10^{-2}$ \\
  & P-MISE & 0.105 & 0.103 & 0.103 & 0.105 \\\hline
\multirow{3}{*}{35} & MISE & 0.384 & 0.383 & 0.384 & 0.386 \\
 & Variance & $7.0 \times 10^{-3}$ & $7.7 \times 10^{-3}$ & $9.3 \times 10^{-3}$ & $1.1 \times 10^{-2}$ \\
  & P-MISE & 0.273 & 0.270 & 0.269, & 0.269 \\\hline
\multirow{3}{*}{37} & MISE & 0.589 & 0.588 & 0.589 & 0.591 \\
 & Variance & $6.0 \times 10^{-3}$ & $6.5 \times 10^{-3}$ & $7.9 \times 10^{-3}$ & $1.0 \times 10^{-2}$ \\
  & P-MISE & 0.473 & 0.470 & 0.469 & 0.470 \\\hline
\end{tabular}
\caption{Mean-integrated squared error (MISE), variance, and partial MISE as a function of $\Delta^\star$ and sampling rate when true $\Delta = 32$ minutes.}
\label{tab:mise_delta}
\end{table}

@@

\begin{figure}[!th]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{../figures/smoothed_eda.png}
  \caption{Electrodermal activity (EDA)}
  \label{fig:mean_eda}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{../figures/smoothed_ai.png}
  \caption{Activity Index (AI)
}  \label{fig:mean_acc}
\end{subfigure}
\caption{Average scaled EDA and AI in the 30 minutes prior to button presses}
\label{fig:mean_edacc}
\end{figure}

@@

% \begin{figure}[!th]
% \centering
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{../figures/missing_data_eda.png}
%   \caption{Electrodermal activity (EDA)}
%   \label{fig:missing_eda}
% \end{subfigure}%
% \begin{subfigure}{.5\textwidth}
%   \centering
%   \includegraphics[width=.8\linewidth]{../figures/missing_data_ai.png}
%   \caption{Activity Index (AI)
% }  \label{fig:missing_acc}
% \end{subfigure}
% \caption{Patient average fraction missingness in 30-minute windows.}
% \label{fig:mean_edacc2}
% \end{figure}
